\emph{Light Exposure Behavior Assessment (LEBA)}: Development of a novel instrument to capture light exposure-related behaviours

### 2021-11-13 manuscript.Rmd
* Longtables are visible

```{r FlowchartFig, echo=FALSE, fig.align='center', fig.cap='Development of long and short form of LEBA', out.height='100%', out.width='100%'}
knitr::include_graphics('Manuscript.figures/Flowchart1.png')
```

### Data Collection
Timeline of data collection, mode of data collection.

## Analytic Strategies
For our analysis we used R (version 4.1.0), with several R packages. Initially, our tool had six point Likert type response scale (0:Does not apply/I don't know; 1:Never, 2:Rarely; 3:Sometimes; 4:Often; 5: Always). Our purpose was to capture light exposure related behavior and these two response options: "Does not apply/I don't know" and "Never" were providing similar information. As such we decided to collapse them into one making it a 5 point Likert type response scale. Necessary assumptions of EFA, including sample adequacy, normality assumptions, quality of correlation matrix, were assessed. Our data violated both the univariate and multivariate normality assumptions. Due to these violations and the ordinal nature of our response data, we used  polychoric correlation matrix [@desjardinsHandbookEducationalMeasurement2018] for the EFA. We employed principal axis (PA) as  factor extraction method with varimax rotation. PA is robust to the normality assumption violations [@watkinsStepbyStepGuideExploratory2020]. The obtained latent structure was confirmed by another factor extraction method: "the minimum residuals extraction" as well. We used a combination of factor identification method including scree plot [@cattellScreeTestNumber1966], Horn's parallel analysis [@hornRationaleTestNumber1965], minimum average partials method [@velicerDeterminingNumberComponents1976], and hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers. Additionally, to determine the simple structure, we followed the following guidelines recommended by psychometricians (i) no factors with fewer than three items (ii) no factors with a factor loading <0.3 (iii) no items with cross-loading greater than .3 across factors [@bandalosFactorAnalysisExploratory2018]. 
We also conducted psychometric analysis on non-merged response options data (supplementary analysis) and rejected the latent structure obtained as the factors were less interpretable.

@@

Sampling adequacy was checked using Kaiser-Meyer-Olkin (KMO) measures of sampling adequacy [@kaiserIndexFactorialSimplicity1974] . The overall KMO vale for 48 items was `r printnum(KMO$MSA)` which was above the cutoff value (.50) indicating a mediocre sample [@hutchesonMultivariateSocialScientist1999].  Bartlett's test of sphericity [@bartlettNoteMultiplyingFactors1954], $\chi^2$ (`r bartlet$df`) = `r apa(bartlet$chisq, 2, T)`, p < .001] indicated the correlations between items are adequate for the EFA. However only `r apa(cor.per, 2,T)`% of the inter-item correlation coefficients were greater than .30. The inter item correlation ranged between `r apa(min.cor,2,F)` to `r apa(max.cor,2,F)`. And the corrected item-total correlations ranged between `r apa(low.r.corec,2,F)` to `r apa(high.r.corec,2,F)`.

```{r tabDes,results='asis'}
apa_table(Des.combined, caption = "Descriptive Statistics", align = "c", note = "*p<.001", longtable =T)
```

@@

```{r MAPefa, include=FALSE}
#MAP
psych::VSS(EFA.data, rotate = "varimax", fm = 'minres', n.obs =428 )
```
Scree plot ( Figure \@ref(fig:facIdFig)) suggested a six-factor solution. Horn's parallel analysis [@hornRationaleTestNumber1965] with 500 iterations also indicated a six-factor solution. However, the minimum average partial (MAP) method [@velicerDeterminingNumberComponents1976] and Hull method [@lorenzo-sevaHullMethodSelecting2011] suggested a five-factor solution. As a result, we tested both five-factor and six-factor solutions.

@@

```{r EFAResiduals, out.height= "50%",  fig.align='center', fig.cap= ' Histogram of residuals:  five-factor solution',  out.width="50%",  warning=FALSE, results='asis'}
hist(residual.5F,  main = NULL, xlab = "Five-factor solution", ylim=c(0, 400))
abline(h =51, lwd = 2, lty = 2)
abline(v =.05,lwd = 2, lty = 2)

@@

```{r itmeinfo, fig.cap="Item information curves (a) Wearing blue light filters (b) Spending time outdoors (c)Using phone and smartwatchin bed (d)Using light before bedtime (e)Using light in the morning andduring daytime", warning=FALSE, out.width="100%", out.height="120%"}
plot_grid(F1_iteminfo, F2_iteminfo,F3_iteminfo,F4_iteminfo,F5_iteminfo, 
          labels = "AUTO", label_size = 10,align = "v")

@@

```{r Testinfo, fig.cap="Test information curves (a) Wearing blue light filters (b) Spending time outdoors (c)Using phone and smartwatchin bed (d)Using light before bedtime (e)Using light in the morning andduring daytime", warning=FALSE, out.width="100%", out.height="120%"}
plot_grid(F1_test.info , F2_test.info.redu,F3_test.info.redu,F4_test.info.redu,F5_test.info,
          labels = "AUTO", label_size = 10, align = "v" )

@@

Table \@ref(tab:itemfittab) summarizes the item fit indexes of the items. All the items fitted well to the respective models as assessed by  RMSEA value obtained from Signed-$\chi^2$ index implementation. All of the items had RMSEA value <.06 indicating adequate fit. Figure \@ref(fig:persornfit) depicts the person fit of out fitted models. Person fit indicates the validity and meaningfulness of the fitted model at the participants latent trait level [@desjardinsHandbookEducationalMeasurement2018]. We estimated the person fit statistics using standardized fit index Zh statistics [@drasgow1985appropriateness]. Zh < -2 should be considered as a misfit. Fig indicates that Zh is larger than -2 for most participants, suggesting a good fit of the selected IRT models.
```{r itemfittab, results='asis',warning=F, message=F}
apa_table(itemfit, caption ="Item fit statistics for the fitted models")
```

```{r personfit, fig.cap="Person fit of the five fitted IRT models (a) Wearing blue light filters (b) Spending time outdoors (c)Using phone and smartwatchin bed (d)Using light before bedtime (e)Using light in the morning andduring daytime", warning=FALSE, out.width="100%", out.height="120%"}
plot_grid(F1.person, F2.person,F3.person,F4.person,F5.person, 
          labels = "AUTO", label_size = 10,align = "v")

@@

```{r scalechar, fig.cap="Scale characteristic curve of the five fitted IRT models ((a) Wearing blue light filters (b) Spending time outdoors (c)Using phone and smartwatchin bed (d)Using light before bedtime (e)Using light in the morning andduring daytime", warning=FALSE, out.width="100%", out.height="120%"}

>>>>

```{r FlowchartFig, echo=FALSE, fig.align='center', fig.cap='Development of long and short form of LEBA', out.height='150%', out.width='100%'}
knitr::include_graphics('Manuscript.figures/Flowchart1.png')
```

### Data Collection

Timeline of data collection, mode of data collection.

## Analytic Strategies

For our analysis we used R (version 4.1.0), with several R packages. Initially, our tool had six point Likert type response scale (0:Does not apply/I don't know; 1:Never, 2:Rarely; 3:Sometimes; 4:Often; 5: Always). Our purpose was to capture light exposure related behavior and these two response options: "Does not apply/I don't know" and "Never" were providing similar information. As such we decided to collapse them into one making it a 5 point Likert type response scale. Necessary assumptions of EFA, including sample adequacy, normality assumptions, quality of correlation matrix, were assessed. Our data violated both the univariate and multivariate normality assumptions. Due to these violations and the ordinal nature of our response data, we used  polychoric correlation matrix [@desjardinsHandbookEducationalMeasurement2018] for the EFA. We employed principal axis (PA) as  factor extraction method with varimax rotation. PA is robust to the normality assumption violations [@watkinsStepbyStepGuideExploratory2020]. The obtained latent structure was confirmed by another factor extraction method: "the minimum residuals extraction" as well. We used a combination of factor identification method including scree plot [@cattellScreeTestNumber1966], Horn's parallel analysis [@hornRationaleTestNumber1965], minimum average partials method [@velicerDeterminingNumberComponents1976], and hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers. Additionally, to determine the simple structure, we followed the following guidelines recommended by psychometricians (i) no factors with fewer than three items (ii) no factors with a factor loading <0.3 (iii) no items with cross-loading greater than .3 across factors [@bandalosFactorAnalysisExploratory2018]. 
We also conducted psychometric analysis on non-merged response options data (supplementary analysis) and rejected the latent structure obtained as the factors were less interpretable.

@@

Sampling adequacy was checked using Kaiser-Meyer-Olkin (KMO) measures of sampling adequacy [@kaiserIndexFactorialSimplicity1974] . The overall KMO vale for 48 items was `r printnum(KMO$MSA)` which was above the cutoff value (.50) indicating a mediocre sample [@hutchesonMultivariateSocialScientist1999].  Bartlett's test of sphericity [@bartlettNoteMultiplyingFactors1954], $\chi^2$ (`r bartlet$df`) = `r apa(bartlet$chisq, 2, T)`, p < .001] indicated the correlations between items are adequate for the EFA. However only `r apa(cor.per, 2,T)`% of the inter-item correlation coefficients were greater than .30. The inter item correlation ranged between `r apa(min.cor,2,F)` to `r apa(max.cor,2,F)`. And the corrected item-total correlations ranged between `r apa(low.r.corec,2,F)` to `r apa(high.r.corec,2,F)`.
```{r tabDes,results='asis'}
apa_table(Des.combined, caption = "Descriptive Statistics", align = "c", note = "*p<.001", longtable =T)
```

@@

```{r MAPefa, include=FALSE}
#MAP
psych::VSS(EFA.data, rotate = "varimax", fm = 'minres', n.obs =428 )
```

Scree plot ( Figure \@ref(fig:facIdFig)) suggested a six-factor solution. Horn's parallel analysis [@hornRationaleTestNumber1965] with 500 iterations also indicated a six-factor solution. However, the minimum average partial (MAP) method [@velicerDeterminingNumberComponents1976] and Hull method [@lorenzo-sevaHullMethodSelecting2011] suggested a five-factor solution. As a result, we tested both five-factor and six-factor solutions.

@@

```{r EFAResiduals, out.height= "100%",  fig.align='center', fig.cap= ' Histogram of residuals:  five-factor solution',  out.width="100%",  warning=FALSE, results='asis'}
hist(residual.5F,  main = NULL, xlab = "Five-factor solution", ylim=c(0, 400))
abline(h =51, lwd = 2, lty = 2)
abline(v =.05,lwd = 2, lty = 2)

@@

```{r itmeinfo, fig.cap="Item information curves (a) Wearing blue light filters (b) Spending time outdoors (c) Using phone and smartwatchin bed (d) Using light before bedtime (e) Using light in the morning andduring daytime", warning=FALSE, out.width="100%", out.height="120%"}
plot_grid(F1_iteminfo, F2_iteminfo,F3_iteminfo,F4_iteminfo,F5_iteminfo, 
          labels = "AUTO", label_size = 10,align = "v")

@@

```{r Testinfo, fig.cap="Test information curves (a) Wearing blue light filters (b) Spending time outdoors (c) Using phone and smartwatchin bed (d) Using light before bedtime (e) Using light in the morning andduring daytime", warning=FALSE, out.width="100%", out.height="120%"}
plot_grid(F1_test.info , F2_test.info.redu,F3_test.info.redu,F4_test.info.redu,F5_test.info,
          labels = "AUTO", label_size = 10, align = "v" )

@@

Table \@ref(tab:itemfittab) summarizes the item fit indexes of the items. All the items fitted well to the respective models as assessed by  RMSEA value obtained from Signed-$\chi^2$ index implementation. All of the items had RMSEA value <.06 indicating adequate fit. Figure \@ref(fig:personfit) depicts the person fit of out fitted models. Person fit indicates the validity and meaningfulness of the fitted model at the participants latent trait level [@desjardinsHandbookEducationalMeasurement2018]. We estimated the person fit statistics using standardized fit index Zh statistics [@drasgow1985appropriateness]. Zh < -2 should be considered as a misfit. Fig indicates that Zh is larger than -2 for most participants, suggesting a good fit of the selected IRT models.
```{r itemfittab, results='asis',warning=F, message=F}
apa_table(itemfit, caption ="Item fit statistics for the fitted models")
```

```{r personfit, fig.cap="Person fit of the five fitted IRT models (a) Wearing blue light filters (b) Spending time outdoors (c) Using phone and smartwatchin bed (d) Using light before bedtime (e) Using light in the morning andduring daytime", warning=FALSE, out.width="100%", out.height="120%"}
plot_grid(F1.person, F2.person,F3.person,F4.person,F5.person, 
          labels = "AUTO", label_size = 10,align = "v")

@@

```{r scalechar, fig.cap="Scale characteristic curve of the five fitted IRT models (a) Wearing blue light filters (b) Spending time outdoors (c) Using phone and smartwatchin bed (d) Using light before bedtime (e) Using light in the morning andduring daytime", warning=FALSE, out.width="100%", out.height="120%"}
